\begin{thebibliography}{99} 

\bibitem{transport_paper}
Nico Leumer, Milena Grifoni, Bhaskaran Muralidharan, and Magdalena Marganska.
 \href{https://doi.org/10.1103/PhysRevB.103.165432}{\textit{Linear and nonlinear transport across a finite Kitaev chain: An exact analytical study}}, 2021

\bibitem{analysis_paper}
Nico Leumer, Magdalena Marganska, Bhaskaran Muralidharan, and Milena Grifoni.
 \href{https://doi.org/10.1088/1361-648X/ab8bf9}{\textit{Exact eigenvectors and eigenvalues of the finite Kitaev chain and its topological properties}}, 2020

\bibitem{entropy_paper}
Abhisek Kejriwal and Bhaskaran Muralidharan.
 \href{https://doi.org/10.1103/PhysRevB.105.L161403}{\textit{Nonlocal conductance and the detection of Majorana zero modes: Insights from von Neumann entrpy}}, 2022

\bibitem{dephasing_paper}
Chaitrali Duse, Praveen Sriram, Kaveh Gharavi, Jonathon Baugh, and Bhaskaran Muralidharan.
 \href{https://doi.org/10.1088/1361-648X/ac0d16}{\textit{Role of dephasing on the conductance signatures of Majorana zero modes}}, 2021

\bibitem{thesis}
Praveen Sriram and Bhaskaran Muralidharan.
 \href{https://doi.org/10.1103/PhysRevB.100.155431}{\textit{Quantum Transport in Superconductor-Semiconductor Hybrid Nanostructures}}, 2019

\bibitem{qt}
Supriyo Dutta.
 \href{https://www.cambridge.org/core/books/quantum-transport/E96BE74AACD59A03A7D6A7F7DACDFB71}{\textit{Quantum Transport - Atom to Transistor}}.
Cambridge University Press, 2005.

\bibitem{superconductivity} 
James Annett.
 \href{https://webuser.unicas.it/pagliarone/Super.pdf}{\textit{Superconductivity, Superfluids, and Condensates}}. 
Oxford University Press, 2004.

\bibitem{youtube_1}
Supriyo Dutta.
 \href{https://www.youtube.com/watch?v=eg4krA0xH6I}{\textit{Fundamentals of Nanoelectronics - Quantum Transport}}.
Youtube, 2015.

\bibitem{youtube_2}
Andrew Mitchell. 
 \href{https://www.youtube.com/playlist?list=PLotxEOxVaaoKRXdDN-7lI3Y88PaHqyOZL}{\textit{Quantum Condensed Matter Physics}}. 
Youtube, 2022.

\bibitem{youtube_3}
Nicola Spaldin. 
 \href{https://www.youtube.com/playlist?list=PL8n8OkcK9SRZwL24mF7FeOa50Xd9YB-JO}{\textit{Schrödinger's Kittens Productions}}. Youtube, 2021.

\bibitem{ssh}
Navketan Batra and Goutam Sheet. 
 \href{https://link.springer.com/article/10.1007/s12045-020-0995-x}{\textit{Physics with Coffee and Doughnuts: Understanding the Physics Behind Topological Insulators Through Su-Schrieffer-Heeger Model}}.
Resonance, Indian Academy of Sciences, 2020.

\bibitem{bloch}
\href{https://bohr.physics.berkeley.edu/classes/221/s07/notes/blochban.pdf}{Lecture notes on Bloch's theorem}, 2005.

\bibitem{tight-binding}
\href{https://www.physics.rutgers.edu/~eandrei/chengdu/reading/tight-binding.pdf}{Lecture notes on tight binding models}, 2015.

\bibitem{second quantization}
Alexander Altland and Ben Simons.
 \href{https://www.tcm.phy.cam.ac.uk/~bds10/tp3/secqu.pdf}{\textit{Quantum Condensed Matter Field Theory}}, 2023.

\end{thebibliography}